Reformulating phylogenetic mixed models to improve flexibility and speed

### 2018-09-05 main.tex
find citition for scaling factor
temp pushing

\newcommand{\bS}{{\mathbf S}}
\newcommand{\bJ}{{\mathbf J}}
\newcommand{\bB}{{\mathbf B}}
\newcommand{\bBadj}{{\mathbf B}_{\mbox{\tiny adj}}}
\newcommand{\bomega}{{\boldsymbol \omega}}
\newcommand{\bell}{{\boldsymbol \ell}}

An alternative approach is to model the phylogenetic correlation as a \textit{Gaussian process}. 
In particular, suppose that the evolutionary process is a Brownian motion, which means the evolution of a continuous trait is a random walk and daughter species after speciation are all independent.  
In that case, the phylogenetic variability of a particular observation can be written as the sum of the evolutionary changes that occurred on all of the branches in the phylogeny in its past. 
Thus, the evolutionary history for each species can be modeled with a sequence of independent errors with species--branch matrix $\bS$, rather than having to impose a correlation $\bC$.

\subsubsection*{Constructing the species--branch matrix}

@@

\mli{example.Rout}
\begin{align*}
\bS_{id} = \begin{bmatrix}
1 & 1 & 0 & 0 \\ 
1 & 0 & 1 & 0 \\ 
0 & 0 & 0 & 1
\end{bmatrix} , 
\bB = \begin{bmatrix}
L1 \\
L2 \\
L3 \\
L4 
\end{bmatrix}
\\
\bS = \bS_{id} \bB = \begin{bmatrix}
L1 & L2 & 0 & 0 \\
L1 &  0 &L3 & 0 \\
0  &  0 & 0  & L4
\end{bmatrix}
\end{align*}

>>>>

\newcommand{\bS}{{\mathbf S}}
\newcommand{\bJ}{{\mathbf J}}
\newcommand{\bB}{{\mathbf B}}
\newcommand{\bBadj}{{\mathbf B}_{\mbox{\tiny adj}}}
\newcommand{\bomega}{{\boldsymbol \omega}}
\newcommand{\bell}{{\boldsymbol \ell}}
\newcommand{\e}{{ \epsilon}}

An alternative approach is to model the phylogenetic correlation as a \textit{Gaussian process}. 
In particular, suppose that the evolutionary process is a Brownian motion, which means the evolution of a continuous trait is a random walk and daughter species after speciation are all independent.  
In that case, the phylogenetic variability of a particular observation can be written as the sum of the evolutionary changes that occurred on all of the branches in the phylogeny in its past. 
Thus, the evolutionary history for each species can be modeled with a sequence of independent errors with species--branch matrix $\bS$, rather than having to impose a correlation $\bC$.
For example, \mli{ref fig:tree} is a phylogenetic tree with three species, the corresponding $\bS$.

\begin{align*}
\bS = \begin{bmatrix}
\ell_1 \e_1 & \ell_2 \e_2 & 0 & 0 \\
\ell_1 \e_1 &  0 & \ell_3 \e_3 & 0 \\
0  &  0 & 0  & \ell_4 \e_4
\end{bmatrix}
\end{align*}
The error value corresponding to tip 1 is $\ell_1 \e_1 + \ell_2 \e_2$, where $\ell_i$ is the square root of the branch length and the $\ell_i$ are independent, homoscedastic Normal variates.

@@

\begin{align*}
\bS_{id} = \begin{bmatrix}
1 & 1 & 0 & 0 \\ 
1 & 0 & 1 & 0 \\ 
0 & 0 & 0 & 1
\end{bmatrix} , 
\bB = \begin{bmatrix}
\ell_1 \\
\ell_2 \\
\ell_3 \\
\ell_4 
\end{bmatrix}
\end{align*}

\mli{cite ives book for standardizing VCV with the determinate scaling factor}
